---
output: github_document
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "70%",
  warning = FALSE,
  message = FALSE
)
```

# micomposition

<!-- badges: start -->
<!-- badges: end -->

**micomposition** is an R package for visualizing microbiome composition
data from 16S rRNA amplicon sequencing or metagenomic studies. It provides
publication-ready plots with minimal code.

## Installation
```r
# Install from GitHub
# install.packages("remotes")
remotes::install_github("YINGSAOVO/micomposition")
```

## Quick Start
```{r load}
library(micomposition)

# Built-in demo dataset
data(demo_data)
str(demo_data, max.level = 1)
```

## Stacked Bar Chart
```{r bar, fig.height=4}
plot_bar_stack(demo_data, level = "Phylum")
```

## Sunburst Chart
```{r sunburst, fig.height=5}
plot_sunburst(demo_data, levels = c("Phylum", "Family"))
```

## Treemap
```{r treemap, fig.height=5}
plot_treemap(demo_data, level = "Family")
```

## Waffle Chart
```{r waffle, fig.height=4}
plot_waffle(demo_data, level = "Phylum")
```

## Radial Bar Chart
```{r radial, fig.height=5}
plot_radial_bar(demo_data, level = "Phylum")
```

## Ridge Plot
```{r ridge, fig.height=6}
plot_ridge(demo_data, level = "Family", group_by = "Group")
```

## Bubble Chart
```{r bubble, fig.height=5}
plot_voronoi_static(demo_data, level = "Family")
```

## Input Data Format

All functions accept a named list with three elements:
```r
data <- list(
  otu_table   = otu_rel,    # numeric matrix: samples × OTUs (relative abundance)
  tax_table   = tax_df,     # data.frame: OTU × taxonomic ranks
  sample_data = sample_df   # data.frame: sample metadata
)
```

Supported input formats (coming soon): `phyloseq`, MetaPhlAn, Kraken/Bracken.

## Citation

If you use micomposition in your research, please cite:

> Your Name (2025). micomposition: Visualization Tools for Microbiome
> Composition Data. R package.
> https://github.com/YINGSAOVO/micomposition
